### 2016-11-21 Projects/DelCo_2012/Report/DelCo_2012.tex
JH edits up to pg. 45 of CAF report

\vspace*{2\baselineskip}
October 2016}

@@

\large
\flushright{\bf National Institute of Standards and Technology Technical Note 1927 \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~1927, \pageref{LastPage} pages (October 2016) \\
CODEN: NTNOEF }

@@

All of the experiments were conducted at full scale in compartments and structures that were representative of room sizes in a residential occupancy. Technical panel members whose departments routinely use CAF for structural fire fighting were invited and present at many of the fire suppression experiments. The experiments were planned during the workshop.  Three types of experiments examining spray density measures, gas layer cooling, and fire suppression were conducted in the fall of 2012. The data from these experiments were presented to the technical panel members. The fire suppression experiments were conducted with a fuel package that included upholstered furniture. After a review of the data, a decision was made to conduct an additional series of suppression experiments with a fuel package that was composed mainly of wood-based materials. The logic was that the synthetic materials in the upholstered furniture burned only on the surface and were easily extinguished. Wood-based fuels would char and present a deep seated fire challenge to the extinguishing agents.  After the wood-based fuel package experiments were completed in 2013, it was decided in discussions with members of the technical panel that other experimental conditions needed to be examined: room size, additional ventilation, and a manually operated handline. The issue of a larger fire compartment had also been raised during the workshop discussion, but the desire for replicate experiments and financial limitations removed the large floor plan experiments from the original plan. To accomplish the last series of experiments, additional funding was obtained from NIST, and the experiments were conducted in 2015. The fuel load contained both wood-based wall and ceiling finishes, pallets, carpeting, and upholstered furniture.

@@

\section{Description of Experiments}
\label{sec:desc_experiments}
The experimental setup for this test series is described by the three types of experiments conducted: spray density measures, gas layer cooling, and fire suppression.  Baseline operating conditions were agreed upon at the CAFS workshop for these experiments (Table~\ref{tab:op_condition})~\cite{Grant:2011}. The nozzles chosen for these experiments were Task Force Tips (TFT) Metro 1, fixed gallonage nozzles with a flow rate of 120~gpm. These are automatic nozzles which have a 7/8~inch (22~mm) slug-tip and can be varied from a straight stream to a wide fog stream for application. The setup also included 100~ft (30.5~m) of 1-3/4~inch (44~mm) hose. Nozzles were mounted to a ``blitz-fire'' monitor device and placed inside the structures. The nozzles were controlled and operated via a gate valve from the exterior. The exterior control had a gated wye to allow the flow of CAF through a hose directed away from the test structure. When the flow of CAF was needed for the experiment, the gated wye would be switched to allow the CAF to flow into the test structure. This was done to minimize the build-up of ``dry'' foam or a slug upon the initial flow of the nozzle into the test structure.

\begin{table}[!ht]
\centering
\captionof{table}{Operating conditions for experiments}\label{tab:op_condition}
\begin{tabular}{ll}
\toprule[1.5pt]
Equipment    &   Operating Condition \\
\midrule
Nozzle       & Metro 1, fixed gallonage fog, straight, or 7/8~inch solid stream \\
Hoseline     & 100~ft of 1-3/4~inch hoseline \\
Foam         & 0.3\% foam concentration \\
Pressure     & Water and Air pressure of 100--120~psi (689--827~kPa)\\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/Flows}
	\caption{Spray Density Nozzle Patterns}
	\label{fig:Spray_Density_Nozzle_Patterns}
\end{figure}

@@

The fire suppression series included 20 experiments. Twelve experiments were conducted in the single-story structures (6 in each structure) using both water and CAF application as a suppression agent. The nozzle patterns were varied and included straight stream, solid stream, and narrow fog stream. The six experiments in the two-story structure used both water and CAF from a straight stream nozzle.  In addition two CAF experiments were conducted in the two-story with a solid stream.

@@

The burn building, or live fire training structure, contained a two-story section and a three-story section. The building was supported with reinforced concrete beams and columns. The floors and ceilings were also concrete, and the interior walls were constructed of cement block. The walls and ceiling of the burn rooms were protected with a 25~mm (1~in) thick layer of calcium silicate insulation and was covered by 50~mm (2~in) thick concrete tile. The floor of the burn rooms was protected with fire brick. Tests were conducted in the two-story, two-room configuration, which had overall dimensions of 6.5~m (21.3~ft) by 9.2~m (30.2~ft) (see Figure~\ref{fig:Delaware_County,_PA_Burn_Building_Layout}).

@@

The floor of the structure was the concrete pad. The interior walls of the structure were framed with steel studs and track. The studs were set to 0.40~m (16~in) centers. The ceiling support was composed of wood truss joist I-beams (TJIs) with a 299~mm (11.75~in) depth. The TJI was composed of laminated veneer lumber flanges with a cross section of 29~mm (1.125~in) x 44~mm (1.75~in) and an 11~mm (0.43~in) thick oriented strand board web. Tongue and grove, 18.3~mm (0.72~in) thick, oriented strand board was attached to the top of the TJIs.

The interior walls of the burn room were lined with 13~mm (0.5~in) thick cement board. The ceiling of the burn room was the exposed ``floor assembly''. The walls of the hallway and entry foyer were composed of 16~mm (0.625~in) Type X gypsum board. The ceiling of the hallway and entry foyer was composed of two layers of 13~mm (0.5~in) thick cement board.

@@

The west structure was modified to add a second story to test larger fires and remote suppression effects (see Figure~\ref{fig:delco_2story}).

@@

The structures were instrumented for gas temperature, gas velocity, and heat flux measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single, sheathed thermocouples were installed in conjunction with the bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple was protected with an 3.2~mm (0.125~in) diameter inconel sheath. The sheathing protects the thermocouple from water but slows the response time because of the additional thermal mass. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. A legend, which clarifies the instrumentation schematics discussed in the following sections, is included in Figure~\ref{fig:Instrumentation_Legend}.

@@

In June 2015, two stacks of pallets were added to the fuel load consisting of three sofas; OSB lining on the ceiling and walls; and the polypropylene carpeting over polyurethane foam padding. Each stack had 10 pallets and an approximate mass of 175~kg (386~lb). Hay from bales averaging a mass of 12~kg (26~lb) was added for ignition purposes.

@@

Spray density experiments were conducted by placing 0.23~m$^2$ (2.5~ft$^2$) interlocking water collection pans across the floor surface area of the burn rooms within each structure (see Figure~\ref{fig:water_buckets}). The depth of the pans was 0.3 m (1 ft).  Water or CAF was flowed for a specified duration or until it appeared that a pan was filled.  Upon completion of the test, the bins were removed from the structure. Then, each bin was weighed to quantify the impact of nozzle location and spray pattern. Thirty-three spray density tests were conducted: 21 in the burn building used for the gas cooling tests and 12 in the single-story structure used for the fire suppression experiments. The interior of the fire room in the single-story structure was sheathed in plastic for the spray density tests to protect the gypsum board from being soaked with water prior to the fire suppression experiments. The flow rate for each test was 120~gpm for water and 120~gpm plus 60~cfm for CAF. Tables~\ref{tab:spray_density_tests} and \ref{tab:spray_density_tests2} provide an overview of each experiment. In the table, ``GC'' stands for the gas cooling experiments, and ``SE'' stands for the suppression experiments. For all experiments, 30~m (100~ft) of 1~3/4~inch hoseline connected to a ``blitz-fire'' monitor with a Metro 1 nozzle was used.

@@

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix for Gas Cooling Experiments}\label{tab:spray_density_tests}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#    &  Duration (s)  & Agent  &  Pattern            & Nozzle Position  \\
\midrule
GC1        &  15            & Water  &  Straight Stream    & Back             \\
GC2        &  23            & Water  &  Straight Stream    & Back             \\
GC3        &  15            & Water  &  Straight Stream    & Mid              \\
GC4        &  15            & Water  &  Straight Stream    & Mid              \\
GC5        &  20            & Water  &  Straight Stream    & Mid              \\
GC6        &  21            & Water  &  30$^{\circ}$ Fog   & Mid              \\
GC7        &  20            & Water  &  30$^{\circ}$ Fog   & Mid              \\
GC8        &  22            & Water  &  30$^{\circ}$ Fog   & Back             \\
GC9        &  24            & Water  &  30$^{\circ}$ Fog   & Back             \\
GC10       &  20            & Water  &  Solid Stream       & Back             \\
GC11       &  22            & Water  &  Solid Stream       & Back             \\
GC12       &  15            & CAF    &  Solid Stream       & Back             \\
GC13       &  10            & CAF    &  Solid Stream       & Mid              \\
GC14       &  15            & CAF    &  Solid Stream       & Back             \\
GC15       &  20            & CAF    &  Solid Stream       & Mid              \\
GC16       &  15            & CAF    &  30$^{\circ}$ Fog   & Mid              \\
GC17       &  20            & CAF    &  Straight Stream    & Mid             \\
GC18       &  10            & CAF    &  Straight Stream    & Mid             \\
GC19       &  15            & CAF    &  30$^{\circ}$ Fog   & Back             \\
GC20       &  15            & CAF    &  30$^{\circ}$ Fog   & Back             \\
GC21       &  15            & CAF    &  Straight Stream    & Back             \\

\bottomrule[1.25pt]
\end{tabular}\par
\end{table}
For the fire suppression spray density experiments, Table~\ref{tab:spray_density_tests2}, the hallway position had the center of the stream aimed at the ceiling of the burn room, 2.14~m (7~ft) north of the south wall and 1.53~m (5~ft) east of the west wall. The fire room position had the center of the stream aimed at the ceiling of the burn room, 2.14~m (7~ft) north of the south wall and 3.05~m (10~ft) west of the east wall.

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix for Fire Suppression Experiments}\label{tab:spray_density_tests2}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#    & Duration (s)  & Agent  &  Pattern            & Nozzle Location  \\
\midrule
SE1        & 7             & Water  &  Straight Stream    &    Fire Room          \\
SE2        & 18            & Water  &  Straight Stream    &    Fire Room          \\
SE3        & 19            & Water  &  Straight Stream    &    Fire Room          \\
SE4        & 21            & Water  &  Straight Stream    &    Fire Room          \\
SE5        & 22            & Water  &  30$^{\circ}$ Fog   &    Fire Room          \\
SE6        & 19            & Water  &  30$^{\circ}$ Fog   &    Fire Room          \\
SE7        & 16            & Water  &  Solid Stream       &    Fire Room          \\
SE8        & 17            & Water  &  Solid Stream       &    Hallway            \\
SE9        & 15            & Water  &  30$^{\circ}$ Fog   &    Hallway            \\
SE10       & 16            & Water  &  30$^{\circ}$ Fog   &    Hallway            \\
SE11       & 16            & Water  &  Straight Stream    &    Hallway            \\
SE12       & 16            & Water  &  Straight Stream    &    Hallway            \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB7}
	\caption{Gas Cooling Spray Density Test 7, 30 $^{\circ}$ Fog} 
	\label{fig:Burn_Building_Test_7}
\end{figure}

@@

There were three thermocouple arrays in the gas cooling room (adjacent to the burn room in Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions}). Figures~\ref{fig:gas_cooling_sub1} and \ref{fig:gas_cooling_sub5} show the variation of the impact of the suppression stream on cooling location for three consecutive water and CAF tests, respectively. For both sets, a solid 7/8~in. nozzle in the middle position was used. The data show that thermocouples from Gas Cooling Array 2 (GC\_2 in the legend), the middle plot of both figures, has more significant temperature decreases than the other two arrays (GC\_1 and GC\_3). This is the result of Gas Cooling Array 2 being directly ``hit'' by the suppression stream.  Gas Cooling Array 1 also exhibits some amount of suppression agent impact.  Gas Cooling Array 3 is located behind the area where the stream impacts the ceiling for both the middle and back nozzle positions.  As a result, it remained dry and seems to provide the best representation of the gas cooling.

@@

To better quantify the effectiveness of the suppression agents (CAF and water), the maximum temperature differences associated with each gas cooling event for each agent from the three thermocouple trees were analyzed. Figure~\ref{fig:combined_all} compares the temperature differences at each thermocouple position (via a scatterplot) for 30 of the 88 experiments. The 30 experiments are actually pairs of experiments, 15 water and 15 CAF, which were directly comparable (see Table~\ref{tab:Gas_Cooling_Descriptions}). These experiments included solid stream, straight streams, and narrow fog streams at both nozzle positions. While the total number of experiments was 88, there were a number of experiments that did not have a directly comparable experiment between water and CAF. For example, a set of 12 Class A solution experiments were conducted. Gas cooling experiments were also conducted with water applied in a wide fog stream. Since no significant differences were noted compared to the narrow fog stream experiments, a comparable series with a wide fog CAF was not conducted. Similarly, larger water flows were examined, but 120 gpm was the maximum water flow that could be generated with the CAFS, so those experiments are also not included in this comparison. The Mitchell report includes the data for all 88 experiments \cite{Mitchell:1}.

\begin{table}[!ht]
\centering
\caption{Gas Cooling Comparison Test Descriptions}\label{tab:Gas_Cooling_Descriptions}
\begin{tabular}{clll}
\toprule[1.5pt]
Water Test $\#$  & CAF Test  $\#$	& Stream			  & Position  \\
\midrule
 2               & 6                &  Solid Stream       &  Middle   \\
 3               & 7                &  Solid Stream       &  Middle   \\
 4               & 10               &  Solid Stream       &  Middle   \\
 9               & 11               &  Solid Stream       &  Middle   \\
 17              & 52               &  Solid Stream       &  Middle   \\
 18              & 53               &  Solid Stream       &  Middle   \\
 19              & 54               &  Solid Stream       &  Middle   \\
 16              & 45               &  Straight Stream    &  Middle   \\
 20              & 46               &  Straight Stream    &  Middle   \\
 21              & 47               &  Straight Stream    &  Middle   \\
 22              & 48               &  Straight Stream    &  Middle   \\
 41              & 50               &  Straight Stream    &  Middle   \\
 29              & 12               &  Straight Stream    &  Full Back   \\
 33              & 13               &  30$^{\circ}$ Fog   &  Full Back   \\
 34              & 14               &  30$^{\circ}$ Fog   &  Full Back   \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

To further analyze the data, the temperature differences can be separated by each of the three thermocouple arrays in the fire room. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2} shows the placement of three thermocouple arrays (TC A1, TC A2, and TC A3). Ideally, the thermocouple is providing an estimate of the temperature of the gas that surrounded it. However, we must remember that the thermocouple is actually providing us with its own temperature. So if a droplet of water or some CAF cling to a thermocouple, the temperature provided by that thermocouple will be different than the surrounding gas temperature.

@@

The thermocouple array data can also be decomposed as a function of nozzle position: 12 middle position experimental pairs and 3 back experimental pairs. Figures~\ref{fig:CAFS_Water_A1_mid}, \ref{fig:CAFS_Water_A2_mid}, and \ref{fig:CAFS_Water_A3_mid} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 12 replicate tests with the monitor nozzle in the middle position. Figures~\ref{fig:CAFS_Water_A1_back}, \ref{fig:CAFS_Water_A2_back}, and \ref{fig:CAFS_Water_A3_back} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 3 replicate tests with the monitor nozzle in the back position. Again, note that the best fit occurs with the data from thermocouple array 3 for both nozzle positions. This analysis of the data results in the same conclusions as the analysis from the Mitchell paper \cite{Mitchell:1}: the gas cooling performance of CAF is similar to water under these conditions.

@@

\begin{figure}[!ht]
	\includegraphics[width=.7\columnwidth]{../Figures/Script_Figures/TCA1_hallnozzle_scatter}
	\caption{CAF vs Water Temperature Difference Comparison for Single-Story Fire Suppression Tests at Thermocouple Array A1 after Hallway Nozzle}
	\label{fig:fs_hall_a1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.7\columnwidth]{../Figures/Script_Figures/TCA1_doornozzle_scatter}
	\caption{CAF vs Water Temperature Difference Comparison for Single-Story Fire Suppression Tests at Thermocouple Array A1 after Door Nozzle}
	\label{fig:fs_door_a1}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=.7\columnwidth]{../Figures/Script_Figures/TCA2_hallnozzle_scatter}
	\caption{CAF vs Water Temperature Difference Comparison for Single-Story Fire Suppression Tests at Thermocouple Array A2 after Hallway Nozzle}
	\label{fig:fs_hall_a2}
\end{figure}

>>>>

\vspace*{2\baselineskip}
December 2016}

@@

\large
\flushright{\bf National Institute of Standards and Technology Technical Note 1927 \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~1927, \pageref{LastPage} pages (December 2016) \\
CODEN: NTNOEF }

@@

All of the experiments were conducted at full scale in compartments and structures that were representative of room sizes in a residential occupancy. Technical panel members whose departments routinely use CAF for structural fire fighting were invited and present at many of the fire suppression experiments. The experiments were planned during the workshop~\cite{Grant:2011}. Three types of experiments examining spray density measures, gas layer cooling, and fire suppression were conducted in the fall of 2012. The data from these experiments were presented to the technical panel members. The fire suppression experiments were conducted with a fuel package that included upholstered furniture. After a review of the data, a decision was made to conduct an additional series of suppression experiments with a fuel package that was composed mainly of wood-based materials. The logic was that the synthetic materials in the upholstered furniture burned only on the surface and were easily extinguished. Wood-based fuels would char and present a deep seated fire challenge to the extinguishing agents.  After the wood-based fuel package experiments were completed in 2013, it was decided in discussions with members of the technical panel that other experimental conditions needed to be examined: room size, additional ventilation, and a manually operated handline. The issue of a larger fire compartment had also been raised during the workshop discussion, but the desire for replicate experiments and financial limitations removed the large floor plan experiments from the original plan. To accomplish the last series of experiments, additional funding was obtained from NIST, and the experiments were conducted in 2015. The fuel load contained both wood-based wall and ceiling finishes, pallets, carpeting, and upholstered furniture.

@@

\section{Description of Experiments}
\label{sec:desc_experiments}
The experimental setup for this test series is described by the three types of experiments conducted: spray density measures, gas layer cooling, and fire suppression.  Baseline operating conditions were agreed upon at the CAFS workshop for these experiments (Table~\ref{tab:op_condition})~\cite{Grant:2011}. The nozzles chosen for these experiments were Task Force Tips (TFT) Metro 1, fixed gallonage nozzles with a flow rate of 120~gpm. These are automatic nozzles which have a 22~mm (0.88~in) slug-tip and can be varied from a straight stream to a wide fog stream for application. The setup also included 30.5~m (100~ft) of 44~mm (1.75~in) diameter hose. Nozzles were mounted to a ``blitz-fire'' monitor device and placed inside the structures. The nozzles were controlled and operated via a gate valve from the exterior. The exterior control had a gated wye to allow the flow of CAF through a hose directed away from the test structure. When the flow of CAF was needed for the experiment, the gated wye would be switched to allow the CAF to flow into the test structure. This was done to minimize the build-up of ``dry'' foam or a slug upon the initial flow of the nozzle into the test structure.

\begin{table}[!ht]
\centering
\captionof{table}{Operating conditions for experiments}\label{tab:op_condition}
\begin{tabular}{ll}
\toprule[1.5pt]
Equipment    &   Operating Condition \\
\midrule
Nozzle       & Metro 1, fixed gallonage fog, straight, or 44~mm (0.88~in) solid stream \\
Hoseline     & 30.5~m (100~ft) of 44~mm (1.75~inch) diameter hoseline \\
Foam         & 0.3\% foam concentration \\
Pressure     & Water and Air pressure of 689--827~kPa (100--120~psi) \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/Flows}
	\caption[Spray Density Nozzle Patterns]{Spray Density Nozzle Patterns (Left-To-Right: Straight Stream, Narrow Fog Stream, and Solid Stream)}
	\label{fig:Spray_Density_Nozzle_Patterns}
\end{figure}

@@

The fire suppression series included 20 experiments. Twelve experiments were conducted in the single-story structures (6 in each structure) using both water and CAF application as a suppression agent. The nozzle patterns were varied and included straight stream, solid stream, and narrow fog stream. The six experiments in the two-story structure used both water and CAF from a straight stream nozzle.  In addition, two CAF experiments were conducted in the two-story burn building with a solid stream.

@@

The burn building, or live fire training structure, contained a two-story section and a three-story section. The building was supported with reinforced concrete beams and columns. The floors and ceilings were also concrete, and the interior walls were constructed of cement block. The walls and ceiling of the burn rooms were protected with a 25~mm (1~in) thick layer of calcium silicate insulation and was covered by 50~mm (2~in) thick concrete tile. The floors of the burn rooms were protected with fire brick. Tests were conducted on the ground floor of the two-story, two-room configuration, which had overall dimensions of 6.5~m (21.3~ft) by 9.2~m (30.2~ft) (see Figure~\ref{fig:Delaware_County,_PA_Burn_Building_Layout}).

@@

The floor of the structure was the concrete pad. The interior walls of the structure were framed with steel studs and track. The studs were set to 0.40~m (16~in) centers. The ceiling support was composed of wood truss joist I-beams (TJIs) with a 299~mm (11.75~in) depth. The TJI was composed of laminated veneer lumber flanges with a cross section of 29~mm (1.13~in) x 44~mm (1.75~in) and an 11~mm (0.43~in) thick oriented strand board web. Tongue and grove, 18.3~mm (0.72~in) thick, oriented strand board was attached to the top of the TJIs.

The interior walls of the burn room were lined with 13~mm (0.5~in) thick cement board. The ceiling of the burn room was the exposed ``floor assembly''. The walls of the hallway and entry foyer were composed of 16~mm (0.63~in) Type X gypsum board. The ceiling of the hallway and entry foyer was composed of two layers of 13~mm (0.5~in) thick cement board.

@@

A second story was added to the west single-story structure in order to test larger fires and remote suppression effects (see Figure~\ref{fig:delco_2story}).

@@

The structures were instrumented for gas temperature, gas velocity, and heat flux measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single, sheathed thermocouples were installed in conjunction with the bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple was protected with an 3.2~mm (0.13~in) diameter inconel sheath. The sheathing protects the thermocouple from water but slows the response time because of the additional thermal mass. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. A legend, which clarifies the instrumentation schematics discussed in the following sections, is included in Figure~\ref{fig:Instrumentation_Legend}.

@@

In June 2015, two stacks of pallets were added to the fuel load consisting of three sofas; OSB lining on the ceiling and walls; and the polypropylene carpeting over polyurethane foam padding. Each stack had 10 pallets and an approximate mass of 175~kg (386~lb). Hay from bales averaging a mass of 12~kg (26~lb) was added for ignition purposes. The specific locations of the fuel load's contents can be found in the dimensioned schematic Figure~\ref{fig:pallet_furniture_2story}.

@@

Spray density experiments were conducted by placing 0.23~m$^2$ (2.5~ft$^2$) interlocking water collection pans across the floor surface area of the burn rooms within each structure (see Figure~\ref{fig:water_buckets}). The depth of the pans was 0.3 m (1 ft).  Water or CAF was flowed for a specified duration or until it appeared that a pan was filled.  Upon completion of the test, the bins were removed from the structure. Then, each bin was weighed to quantify the impact of nozzle location and spray pattern. Thirty-three spray density tests were conducted: 21 in the burn building used for the gas cooling tests and 12 in the single-story structure used for the fire suppression experiments. The interior of the fire room in the single-story structure was sheathed in plastic for the spray density tests to protect the gypsum board from being soaked with water prior to the fire suppression experiments. The flow rate for each test was 120~gpm for water and 120~gpm plus 60~cfm air for CAF. Tables~\ref{tab:spray_density_tests} and \ref{tab:spray_density_tests2} provide an overview of each experiment. In the table, ``GC'' stands for the gas cooling experiments, and ``SE'' stands for the suppression experiments. For all experiments, 30~m (100~ft) of 4.45~cm (1.75~in) diameter hoseline connected to a ``blitz-fire'' monitor with a Metro 1 nozzle was used.

@@

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix for Gas Cooling Experiments}\label{tab:spray_density_tests}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#    &  Duration (s)  & Agent  &  Pattern            	& Nozzle Position  \\
\midrule
GC1        &  15            & Water  &  Straight Stream    	& Back             \\
GC2        &  23            & Water  &  Straight Stream    	& Back             \\
GC3        &  15            & Water  &  Straight Stream    	& Mid              \\
GC4        &  15            & Water  &  Straight Stream    	& Mid              \\
GC5        &  20            & Water  &  Straight Stream 	& Mid              \\
GC6        &  21            & Water  &  Narrow Fog   		& Mid              \\
GC7        &  20            & Water  &  Narrow Fog   		& Mid              \\
GC8        &  22            & Water  &  Narrow Fog   		& Back             \\
GC9        &  24            & Water  &  Narrow Fog   		& Back             \\
GC10       &  20            & Water  &  Solid Stream       	& Back             \\
GC11       &  22            & Water  &  Solid Stream       	& Back             \\
GC12       &  15            & CAF    &  Solid Stream       	& Back             \\
GC13       &  10            & CAF    &  Solid Stream       	& Mid              \\
GC14       &  15            & CAF    &  Solid Stream       	& Back             \\
GC15       &  20            & CAF    &  Solid Stream       	& Mid              \\
GC16       &  15            & CAF    &  Narrow Fog   		& Mid              \\
GC17       &  20            & CAF    &  Straight Stream    	& Mid             \\
GC18       &  10            & CAF    &  Straight Stream    	& Mid             \\
GC19       &  15            & CAF    &  Narrow Fog   		& Back             \\
GC20       &  15            & CAF    &  Narrow Fog   		& Back             \\
GC21       &  15            & CAF    &  Straight Stream    	& Back             \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}
For the fire suppression spray density experiments, Table~\ref{tab:spray_density_tests2}, the hallway position had the center of the stream aimed at the ceiling of the burn room, 2.14~m (7~ft) north of the south wall and 1.53~m (5~ft) east of the west wall. The fire room position had the center of the stream aimed at the ceiling of the burn room, 2.14~m (7~ft) north of the south wall and 3.05~m (10~ft) west of the east wall.

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix for Fire Suppression Experiments}\label{tab:spray_density_tests2}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#    & Duration (s)  & Agent  &  Pattern            & Nozzle Location  	\\
\midrule
SE1        & 7             & Water  &  Straight Stream    &    Fire Room          \\
SE2        & 18            & Water  &  Straight Stream    &    Fire Room          \\
SE3        & 19            & Water  &  Straight Stream    &    Fire Room          \\
SE4        & 21            & Water  &  Straight Stream    &    Fire Room          \\
SE5        & 22            & Water  &  Narrow Fog  		  &    Fire Room          \\
SE6        & 19            & Water  &  Narrow Fog   	  &    Fire Room          \\
SE7        & 16            & Water  &  Solid Stream       &    Fire Room          \\
SE8        & 17            & Water  &  Solid Stream       &    Hallway            \\
SE9        & 15            & Water  &  Narrow Fog   	  &    Hallway            \\
SE10       & 16            & Water  &  Narrow Fog   	  &    Hallway            \\
SE11       & 16            & Water  &  Straight Stream    &    Hallway            \\
SE12       & 16            & Water  &  Straight Stream    &    Hallway            \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB7}
	\caption{Gas Cooling Spray Density Test 7, Narrow Fog} 
	\label{fig:Burn_Building_Test_7}
\end{figure}

@@

There were three thermocouple arrays in the gas cooling room (adjacent to the burn room in Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions}). Figures~\ref{fig:gas_cooling_sub1} and \ref{fig:gas_cooling_sub5} show the variation of the impact of the suppression stream on cooling location for three consecutive water and CAF tests, respectively. For both sets, a solid stream from a 22~mm (0.88~in) nozzle in the middle position was used. The data show that thermocouples from Gas Cooling Array 2 (GC\_2 in the legend), the middle plot of both figures, has more significant temperature decreases than the other two arrays (GC\_1 and GC\_3). This is the result of Gas Cooling Array 2 being directly ``hit'' by the suppression stream.  Gas Cooling Array 1 also exhibits some amount of suppression agent impact.  Gas Cooling Array 3 is located behind the area where the stream impacts the ceiling for both the middle and back nozzle positions.  As a result, it remained dry and seems to provide the best representation of the gas cooling.

@@

To better quantify the effectiveness of the suppression agents, the maximum temperature differences associated with each gas cooling event for each agent from the three thermocouple trees were analyzed. Figure~\ref{fig:combined_all} compares the temperature differences at each thermocouple position (via a scatterplot) for 30 of the 88 experiments. The 30 experiments are actually pairs of experiments, 15 water and 15 CAF, that were directly comparable (see Table~\ref{tab:Gas_Cooling_Descriptions}). These experiments included solid streams, straight streams, and narrow fog streams at both nozzle positions. While the total number of experiments was 88, there were a number of experiments that did not have a directly comparable experiment between water and CAF. For example, 12 of the experiments used Class A solution as the suppression agent. Furthermore, gas cooling experiments were conducted with water applied in a wide fog stream, but no experiments with CAF applied in a wide fog stream were conducted because no significant differences were noted between CAF and water in the narrow fog stream experiments. Similarly, larger water flows were examined, but 120 gpm was the maximum water flow that could be generated with the CAFS. So, those experiments are also not included in this comparison. The Mitchell report includes the data for all 88 experiments \cite{Mitchell:1}.

\begin{table}[!ht]
\centering
\caption{Gas Cooling Comparison Test Descriptions}\label{tab:Gas_Cooling_Descriptions}
\begin{tabular}{clll}
\toprule[1.5pt]
Water Test $\#$  & CAF Test  $\#$	& Stream			  & Position  \\
\midrule
 2               & 6                &  Solid Stream       &  Middle   \\
 3               & 7                &  Solid Stream       &  Middle   \\
 4               & 10               &  Solid Stream       &  Middle   \\
 9               & 11               &  Solid Stream       &  Middle   \\
 17              & 52               &  Solid Stream       &  Middle   \\
 18              & 53               &  Solid Stream       &  Middle   \\
 19              & 54               &  Solid Stream       &  Middle   \\
 16              & 45               &  Straight Stream    &  Middle   \\
 20              & 46               &  Straight Stream    &  Middle   \\
 21              & 47               &  Straight Stream    &  Middle   \\
 22              & 48               &  Straight Stream    &  Middle   \\
 41              & 50               &  Straight Stream    &  Middle   \\
 29              & 12               &  Straight Stream    &  Full Back   \\
 33              & 13               &  Narrow Fog   	  &  Full Back   \\
 34              & 14               &  Narrow Fog   	  &  Full Back   \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

To further analyze the data, the temperature differences can be separated by each of the three thermocouple arrays in the fire room. Figure~\ref{fig:Gas_Cooling_Instrumentation_Dimensions2} shows the placement of three thermocouple arrays (TC A1, TC A2, and TC A3). Ideally, the thermocouple is providing an estimate of the temperature of the gas that surrounded it. However, the thermocouple is actually measuring its own temperature. So, if a droplet of water or some CAF clings to a thermocouple, the temperature provided by that thermocouple will be different than the surrounding gas temperature.

@@

The thermocouple array data can also be decomposed as a function of nozzle position: 12 middle position experimental pairs and 3 back experimental pairs. Figures~\ref{fig:CAFS_Water_A1_mid}, \ref{fig:CAFS_Water_A2_mid}, and \ref{fig:CAFS_Water_A3_mid} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 12 replicate tests with the monitor nozzle in the middle position. Figures~\ref{fig:CAFS_Water_A1_back}, \ref{fig:CAFS_Water_A2_back}, and \ref{fig:CAFS_Water_A3_back} show the temperature difference scatter at thermocouple arrays A1, A2, and A3 respectively for the 3 replicate tests with the monitor nozzle in the back position. Again, note that the best fit occurs with the data from thermocouple array 3 for both nozzle positions. This analysis of the data results in the same conclusions as the analysis from the Mitchell paper \cite{Mitchell:1}: the gas cooling performance of CAF is similar to water under these conditions. Or, more accurately, water is nearly always equal to or better than CAF.

@@

% \begin{figure}[!ht]
% 	\includegraphics[width=.7\columnwidth]{../Figures/Script_Figures/TCA1_hallnozzle_scatter}
% 	\caption{CAF vs Water Temperature Difference Comparison for Single-Story Fire Suppression Tests at Thermocouple Array A1 after Hallway Nozzle}
% 	\label{fig:fs_hall_a1}
% \end{figure}

% \begin{figure}[!ht]
% 	\includegraphics[width=.7\columnwidth]{../Figures/Script_Figures/TCA1_doornozzle_scatter}
% 	\caption{CAF vs Water Temperature Difference Comparison for Single-Story Fire Suppression Tests at Thermocouple Array A1 after Door Nozzle}
% 	\label{fig:fs_door_a1}
% \end{figure}

@@

% \begin{figure}[!ht]
% 	\includegraphics[width=.7\columnwidth]{../Figures/Script_Figures/TCA2_hallnozzle_scatter}
% 	\caption{CAF vs Water Temperature Difference Comparison for Single-Story Fire Suppression Tests at Thermocouple Array A2 after Hallway Nozzle}
% 	\label{fig:fs_hall_a2}
% \end{figure}
